An Ordered Logic for Category Theory

### 2022-09-02 paper/paper.tex
work on shortened synthetic cats section

\begin{document}
%% \newtheorem{theorem}{Theorem}
%% \newtheorem{lemma}{Lemma}
\newtheorem{construction}{Construction}
%% \newtheorem{definition}{Definition}
\newcommand{\sem}[1]{\llbracket{#1}\rrbracket}
\newcommand{\pto}{\nrightarrow}
\newcommand{\pfrom}{\nleftarrow}
\newcommand{\vcat}{\mathcal}
\newcommand{\cat}{\mathbbm}
\newcommand{\isaSet}{\,\,\textrm{Set}}
\newcommand{\isaTy}{\,\,\textrm{Type}}
\newcommand{\isaCat}{\,\,\textrm{Cat}}
\newcommand{\Set}{\textrm{Set}}
\newcommand{\vtkmnd}{\mathbb{K}\text{Mod} (\vcat{V},T)}
\newcommand{\rmod}{\text{RMod}}
\newcommand{\lmod}{\text{LMod}}
\newcommand{\varr}[2]{\text{Fun}{#1}{#2}}
\newcommand{\harr}[2]{\text{Rel}{#1}{#2}}
\newcommand{\harrapp}[3]{#1(#2;#3)}

@@

\newcommand{\homunary}[4]{\pendallXdotY {#1} {\homrallXYtoZ {#3} {#2} {#4}}}
\newcommand{\hombinary}[6]{\pendallXdotY {#1} {\homrallXYtoZ {#3} {#2} \homrallXYtoZ {#5} {#4} {#6}}}

@@

\newcommand{\Cat}{\textrm{Cat}}
\newcommand{\Id}[3]{\textrm{Id} #1\,#2\,#3}
\newcommand{\vlambda}[1]{\lambda^{F}{#1}.}
\newcommand{\hlambda}[2]{\lambda^{P}{#1};{#2}.}

@@

\newcommand{\paramPresheaf}[1]{\mathcal P^{#1}}
\newcommand{\pmPresheaf}{\paramPresheaf\pm}
\newcommand{\posPresheaf}{\paramPresheaf+}
\newcommand{\negPresheaf}{\paramPresheaf-}
\newcommand{\negPresheafApp}[2]{#1\,#2}
\newcommand{\posPresheafApp}[2]{#1\,#2}

@@

\title{A Simple Type Theory for Category Theory}
\author{Max S. New\inst{1,2} and Daniel R. Licata\inst{2}}
\institute{University of Michigan \and Wesleyan University}

@@

\begin{abstract}
  We present a domain-specific type theory for constructions and
  proofs in category theory. The type theory axiomatizes notions of
  category, functor, profunctor and a generalized form of natural
  transformations. By restricting the type theory to a form of ordered
  higher-order logic, we guarantee that all mappings between categories
  are functorial and all transformations are natural by construction,
  with no separate proof necessary. Important category theoretic
  proofs such as the Yoneda lemma and Co-yoneda lemma become simple
  type theoretic proofs about the relationship between unit, tensor
  and (ordered) function types.

  Finally, we give a sound and complete notion of model of our type
  theory in a structure we call a hyperdoctrine of pro-arrow
  equipments with displays, a generalization of prior notions of
  virtual equipments that have been shown to model enriched and
  internal category theory. This means that while the proofs in our
  type theory look like standard set-based arguments, the
  substructural syntactic restrictions ensure that the proofs carry
  over to enriched and internal settings as well.
\end{abstract}

@@

@@

\begin{figure}
  \begin{center}
    \begin{tabular}{ |c|c| } 
      \hline
      Higher-Order Logic & Category Logic \\
      \hline
      Set $X$ & Category $\mathbb{C}$\\
      $X \times Y$ & $\mathbb C \times \mathbb D$ \\ 
      1 & $1$ \\
      $\mathcal P X$ & $\mathcal P^+ X, \mathcal P^- X$ \\
      $\{ (x,y) \in X \times Y | R(x,y) \}$ & $\sum_{\alpha:C;\beta:D} R$\\
      Function $f(x:X) : Y$ & Functor $F(\alpha:\cat C): \cat B$\\
      Relation $R(x,y)$ & Profunctor $R(\alpha:\cat C; \beta: \cat D)$\\
      $R \wedge Q$ & $R \times Q$\\
      $\top$ & $1$\\
      $\forall x. P \Rightarrow Q$ & $\homrallXYtoZ {\alpha:\cat C} P Q, \homlallXYtoZ {\alpha:\cat C} P Q$ \\
      $\exists x. P \wedge Q$ & $\tensorexistsXwithYandZ {\alpha:\cat C} P Q$\\
      $x =_X y$ & $\punitinXfromYtoZ {\cat C} \alpha \beta$\\
      Proof $R_1 \wedge \cdots \Rightarrow Q$ & Natural Transformation $x_1: R_1,\ldots \vdash t : Q$\\
      \hline
    \end{tabular}
  \end{center}
  \caption{Analogy between Higher-Order Logic and Category Logic
    Judgments and Connectives}
\end{figure}

@@

\section{Synthetic Category Theory}

One common source of category theoretic arguments is the duality
principle: since the opposite of any category is also a category, any
statement that quantifies over all categories is equally true of their
opposites. Furthermore in a formal language of categories, any
statement can be dualized to a statement where domains and codomains
are swapped and composition is flipped.
%
A similar duality principle holds in SCT.
%
The basic idea is to reverse the orientation of all \emph{set}
expressions by swapping which variable is covariant and which is
contravariant.
%
This has the effect of reversing the order of the context $\Phi$.

We can define this as a translation $\cdot^* : SCT \to SCT$ defined on
syntactic forms of SCT.

\begin{figure}
    \begin{align*}
      (\harr P Q)^* &= \harr {P^*} {Q^*}\\
      (\alpha:\cat C)^* &= \alpha: \cat C^*\\
      (\alpha:\cat C; \beta:\cat D)^* &= \beta:\cat D^*; \alpha:\cat C^*\\
      (\Phi,x:R,\Psi)^* &= \Psi^*,x:R^*, \Phi^*\\
      (\homrallXYtoZ \alpha P Q)^* &= \homlallXYtoZ \alpha {P^*} {Q^*}\\
      (\homlallXYtoZ \alpha P Q)^* &= \homrallXYtoZ \alpha {P^*} {Q^*}\\
      (\homrlambdaXatYdotZ x \alpha s)^* &= \homllambdaXatYdotZ x \alpha s^*\\
      (\homrappXtoYatZ s t a)^* &= \homlappXtoYatZ {s^*} {t^*} {a^*}\\
      (\tensorexistsXwithYandZ \beta P Q)^* &=
      \tensorexistsXwithYandZ \beta {Q^*}{P^*}\\
      (\tensorintroatXwithYandZ b s t)^* &= \tensorintroatXwithYandZ {b^*} {t^*} {s^*}\\
      (\punitinXfromYtoZ {\cat C} {a_1}{a_2})^* &= \punitinXfromYtoZ {\cat C^*} {a_1^*}{a_2^*}
    \end{align*}

  \caption{Duality Involution}
\end{figure}

@@

>>>>

\begin{document}
%% \newtheorem{theorem}{Theorem}
%% \newtheorem{lemma}{Lemma}
\newtheorem{construction}{Construction}
%% \newtheorem{definition}{Definition}
\newcommand{\ohol}{OHol}
\newcommand{\sem}[1]{\llbracket{#1}\rrbracket}
\newcommand{\pto}{\nrightarrow}
\newcommand{\pfrom}{\nleftarrow}
\newcommand{\vcat}{\mathcal}
\newcommand{\cat}{\mathbbm}
\newcommand{\isaSet}{\,\,\textrm{Set}}
\newcommand{\isaTy}{\,\,\textrm{Type}}
\newcommand{\isaCat}{\,\,\textrm{Cat}}
\newcommand{\Set}{\textrm{Set}}
\newcommand{\vtkmnd}{\mathbb{K}\text{Mod} (\vcat{V},T)}
\newcommand{\rmod}{\text{RMod}}
\newcommand{\lmod}{\text{LMod}}
\newcommand{\varr}[2]{\text{Fun}{#1}{#2}}
\newcommand{\harr}[2]{\text{Rel}{#1}{#2}}
\newcommand{\harrapp}[3]{#1(#2;#3)}

@@

\newcommand{\homunary}[4]{\pendallXdotY {#1} {\homrallXYtoZ {#3} {#2} {#4}}}
\newcommand{\hombinary}[6]{\pendallXdotY {#1} {\homrallXYtoZ {#3} {#2} \homrallXYtoZ {#5} {#4} {#6}}}
\newcommand{\homtrinary}[8]{\pendallXdotY {#1} {\homrallXYtoZ {#3} {#2} \homrallXYtoZ {#5} {#4} \homrallXYtoZ {#7} {#6} {#8}}}

\newcommand{\homlunary}[4]{\pendallXdotY {#1} {\homrallXYtoZ {#3} {#2} {#4}}}
\newcommand{\homlbinary}[6]{\pendallXdotY {#1} {\homrallXYtoZ {#3} {#2} \homrallXYtoZ {#5} {#4} {#6}}}
\newcommand{\homltrinary}[8]{\pendallXdotY {#1} {\homrallXYtoZ {#3} {#2} \homrallXYtoZ {#5} {#4} \homrallXYtoZ {#7} {#6} {#8}}}

@@

\newcommand{\Cat}{\textrm{Cat}}
\newcommand{\Id}[3]{\textrm{Id} #1\,#2\,#3}
\newcommand{\vlambda}[1]{\lambda^{F}{#1}.}
\newcommand{\hlambda}[2]{\lambda^{P}{#1};{#2}.}
\newcommand{\ran}[1]{\textrm{Ran}_{#1}\,}

@@

\newcommand{\paramPresheaf}[1]{\mathcal P^{#1}}
\newcommand{\pmPresheaf}{\paramPresheaf\pm}
\newcommand{\posPresheaf}{\paramPresheaf+}
\newcommand{\negPresheaf}{\paramPresheaf-}
\newcommand{\negPresheafApp}[2]{#1\,#2}
\newcommand{\negPresheafAppPtoX}[2]{#2\,#1}
\newcommand{\posPresheafApp}[2]{#1\,#2}

@@

\title{An Ordered Logic for Category Theory}
\author{Max S. New\inst{1,2} and Daniel R. Licata\inst{2}}
\institute{University of Michigan \and Wesleyan University}

@@

\begin{abstract}
  We present a domain-specific type theory for constructions and
  proofs in category theory. The type theory axiomatizes notions of
  category, functor, profunctor and a generalized form of natural
  transformations. By restricting the type theory to an ordered
  linear restriction of higher-order logic, we guarantee that all
  mappings between categories are functorial and all transformations
  are natural by construction, with no separate proof
  necessary. Important category theoretic proofs such as the Yoneda
  lemma and Co-yoneda lemma become simple type theoretic proofs about
  the relationship between unit, tensor and (ordered) function types,
  and can be seen to be ordered refinements of theorems in
  intuitionistic higher-order logic.

  We give a sound and complete notion of model of our type theory in a
  structure we call a hyperdoctrine of pro-arrow equipments with
  displays, which are known to model enriched and internal category
  theory. This means that while the proofs in our type theory look
  like standard set-based arguments, the substructural syntactic
  restrictions ensure that the proofs carry over to enriched and
  internal settings as well.
\end{abstract}

@@

\subsection{Ordered Logic of Functions and (Constructive) Relations}

Our logic is based on four basic notions: categories, functors,
profunctors and (multi-argument) natural transformations.
%
These generalize the basic notions already present in formulations of
higher-order logic: sorts, functions, relations and proofs.
%
In fact, one way to view our logic is that it is an \emph{ordered},
\emph{constructive} higher-order logic

While profunctors are not typically employed in introductory category
theory texts,

@@

@@

\section{Synthetic Category Theory in Ordered Higher-order Logic}

To demonstrate that \ohol{} provides a setting in which to carry out
synthetic category theory, we demonstrate some basic definitions and
theorems in category theory in \ohol{}. We emphasize throughout the
interpretations as tautologies in higher order-logic, as well as the
relation to ordered linear logic.

First, we fix some terminology: TODO, isomorphism of profunctors, anything else we'll need?
%% \begin{definition}{Isomorphism of Profunctors}

%% \end{definition}

We can witness the fact that all constructions are (pro-)functorial by
using the hom type $\punitinXfromYtoZ {\cat C}{\alpha}{\alpha'}$.
\begin{lemma}
  \begin{enumerate}
  \item Identity: $\pendallXdotY {\alpha} {\punitinXfromYtoZ {\cat C}{\alpha}{\alpha'}}$
  \item Composition: $\hombinary {\alpha_1} {\punitinXfromYtoZ {}{\alpha_1}{\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {}{\alpha_1}{\alpha_2}} {\alpha_3} {\punitinXfromYtoZ {}{\alpha_1}{\alpha_3}}$
  \item Functoriality: for any $F : \varr {\cat C}{\cat D}$, $\homunary {\alpha_1} {\punitinXfromYtoZ {}{\alpha_1}{\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {} {F(\alpha_1)}{F(\alpha_2)}}$.
   \item Profunctoriality: for any $R : \harr {\cat C}{\cat D}$ $\homtrinary {\alpha_1} {\punitinXfromYtoZ{} {\alpha_1}{\alpha_2}} {\alpha_2} {R{\alpha_2}{\beta_2}} {\beta_2} {\punitinXfromYtoZ{} {\beta_2} {\beta_1}} {\beta_1} {R\alpha_1\beta_1}$
  \end{enumerate}
\end{lemma}
Reflexivity and transitivity generalize the corresponding properties
of equality, with the lack of symmetry being a key feature of the
generalization.
%
In addition, we can prove that the (pro)-functoriality axioms commute
with the composition proof by the $\eta$ principle for the unit.
%
(Pro-)Functoriality generalizes the statement that all functions and
relations respect equality.

Naturality is more complex to state, and it is a statement about the
\emph{proofs} so it has no analogue in ordinary higher-order
logic. The following version is stated for any \emph{profunctor}, with
the usual case of naturality arising when $R \alpha \beta =
\punitinXfromYtoZ {\cat C}{F \alpha}{G \beta}$.
\begin{lemma}[Naturality]
  For any $t : \pendallXdotY {\alpha:\cat C}{\harrapp R \alpha
    \alpha}$, by composing with profunctoriality, we can construct
  terms
  \[ \alpha_1, f : \punitinXfromYtoZ {\cat C} {\alpha_1} {\alpha_2}, \alpha_2 \vdash \textrm{lcomp}(f,\pendappXtoY t {\alpha_2}) \quad \textrm{and} \quad \textrm{rcomp}(\pendappXtoY t {\alpha_1}, f) : \harrapp R {\alpha_1}{\alpha_2}\]
  These are both equal to $\punitelimtoYwithkontZ f {t}$.
\end{lemma}

Next, we turn to some of the central theorems of category theory, the
Yoneda and Co-Yoneda lemmas. Despite being ulimately quite elementary,
these are notoriously abstract. In \ohol{}, we can view these as
ordered generalizations of some very simple tautologies about
equality. For instance, the Yoneda lemma (and its proof) generalizes
the equivalence between the formulae $\forall x. \forall y. x = y
\Rightarrow P y$ and $P x$.
\begin{lemma}
  Let $\alpha^o : \cat C$ and $\pi : \negPresheaf {\cat C}$. Then
  \begin{enumerate}
  \item (Yoneda) The profunctor $\homrallXYtoZ {\alpha'} {\punitinXfromYtoZ {\cat C} {\alpha'}{\alpha}} {\negPresheafAppPtoX \pi {\alpha'}}$ is isomorphic to $\negPresheafAppPtoX \pi \alpha$
  \item (Co-Yoneda) The profunctor $\tensorexistsXwithYandZ {\alpha'}{\punitinXfromYtoZ{}{\alpha}{\alpha'}} {\negPresheafAppPtoX \pi \alpha}$ is isomorphic to $\negPresheafAppPtoX \pi \alpha$
  \end{enumerate}
\end{lemma}

Next, we have the ``Fubini'' theorems. These correspond to quantifier
re-ordering and associativity of conjunction and implication in
higher-order logic.
\emph{Fubini} theorems about interchange of (co)-ends.
\begin{lemma}[{Fubini}]
  \label{lem:fubini}
  The following isomorphisms hold when the corresponding profunctors
  are well typed.
  \begin{enumerate}
  \item $\tensorexistsXwithYandZ {\beta} {\harrapp P \alpha \beta} {(\tensorexistsXwithYandZ \gamma {\harrapp Q \beta \gamma} {\harrapp R \gamma \delta})} \cong \tensorexistsXwithYandZ {\gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\harrapp R \gamma \delta}$
  \item $\homrallXYtoZ {\gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\harrapp S \alpha \gamma} \cong \homrallXYtoZ \beta {\harrapp P \alpha \beta} {\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {\harrapp S \alpha \gamma}}$
  \item $\homlallXYtoZ {\alpha} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\harrapp S \alpha \gamma} \cong \homlallXYtoZ \beta {\harrapp Q \beta \gamma} {\homlallXYtoZ \alpha {\harrapp P \alpha \beta} {\harrapp S \alpha \gamma}}$
  \item $\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {(\homlallXYtoZ \alpha {\harrapp P \alpha \beta} {\harrapp S \alpha \gamma})} \cong \homlallXYtoZ \alpha {\harrapp P \alpha \beta} {(\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {\harrapp S \alpha \gamma})}$
  \item $\pendallXdotY \alpha \homrallXYtoZ \beta P {P'} \cong \pendallXdotY \beta \homlallXYtoZ \alpha P {P'}$
  \end{enumerate}  
\end{lemma}
Note that the last item here is different from the others in that it
is an equivalence of sets rather than profunctors.

Next, we turn to basic universal properties. While there are many
equivalent definitions of a universal property in category theory, the
one most natural in \ohol{} is the notion of a functor representing a
profunctor, a mild generalization of the more common notion of
representable presheaf. We say a functor $F : \varr {\cat C}{\cat D}$
left represents a profunctor $R : \harr {\cat C}{\cat D}$ when
$\punitinXfromYtoZ {}{\beta}{F \alpha} \cong \harrapp R \alpha
\beta$. This is an ordered generalization of the notion of a relation
$R$ being the \emph{graph} of a function $F$. Due to the asymmetry,
there is a dual notion of a functor $G : \varr {\cat D}{\cat C}$
right-representing a profunctor $\punitinXfromYtoZ {\alpha}{G\beta}
\cong \harrapp R \alpha \beta$. Then $F$ and $G$ are adjoint functors
when $F$ left represents a profunctor that $G$ right-represents. This
generalizes the formulation that a function $g$ is an inverse to $f$
if its corresponding relation is the converse of the relation
corresponding to $f$.

The definition we gave for adjoint functors is the one in terms of
``hom-isomorphism''. The equivalent in terms of unit and co-unit is a
simple corollary of the Yoneda lemma, because for instance the unit
can be extracted from the transformation $\homunary \alpha
{\punitinXfromYtoZ {}{F \alpha}{\beta}} \beta {\punitinXfromYtoZ
  {}{\alpha}{G\beta}}$. The formulation in terms of unit and co-unit
generalizes the formulation of inverses that each composition is equal
to the identity.
%% \begin{lemma}[Formulations of Representability]
%%   The following are equivalent
%%   \begin{enumerate}
%%   \item An isomorphism between $\punitinXfromYtoZ {}{F \alpha}{\beta}$ and $\harrapp R \alpha \beta$
%%   \item A term $\pendallXdotY \alpha {\harrapp R \alpha {F \alpha}}$
%%     such that the induced morphism from $\punitinXfromYtoZ {}{F
%%       \alpha}{\beta}$ to $\harrapp R \alpha \beta$ is an isomorphism.
%%   \end{enumerate}
%% \end{lemma}

Next, we turn to limits, or more generally, Kan extensions.
%
Given functors $J : \varr {\cat C}{\cat C'}$ and $F : \varr{\cat
  C}{\cat D}$, a right Kan extension is functor $\ran J F : \varr{\cat
  C'}{\cat D}$ such that the composition $\ran J F \circ J$ has a
universal natural transformation to $F$.
\begin{lemma}
  A right Kan extension of $F$ along $J$ is a functor $\ran J F$ that
  right-represents the profunctor $\homrallXYtoZ {\alpha} {\punitinXfromYtoZ {} {\alpha'}{J \alpha}}{\punitinXfromYtoZ {} \beta {F \alpha}}$
\end{lemma}
Viewing this in higher-order logic, this says that $\beta = (\ran J
F)\alpha'$ if and only if $\forall \alpha. \alpha' = J \alpha
\Rightarrow \beta = F \alpha$. If $J$ is surjective this says that
$\ran J F$ composed with $J$ is equal to $F$, but if $J$ is not
surjective then this is impossible to satisfy for ordinary
functions. Kan extensions generalize limits in that the limit of a
diagram $F : \varr {\cat C}{\cat D}$ is the same as a right Kan
extension along the unique $! : \varr{\cat C}{1}$.

Then we can prove the famous theorem that right adjoints preserve
right Kan extensions, and so in particular, limits.
\begin{theorem}
  If $\ran J F$ is a right Kan extension and $G : \varr{\cat D}{\cat
    E}$ has a left adjoint $L$, then $G \circ (\ran J F)$ is the right
  Kan extension of $G \circ F$ along $J$.
\end{theorem}
\begin{proof}
  \begin{align*}
    \punitinXfromYtoZ {} {\gamma} {G (\ran J F \alpha')}
    &\cong \punitinXfromYtoZ {}{L \gamma} {\ran J F \alpha'}\\
    &\cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {}{L \gamma}{F \alpha}}\\
    &\cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {}{\gamma}{G (F \alpha)}}
  \end{align*}  
\end{proof}

\subsection{Ordered Theories}

Just as we can define theories in first-order logic or higher-order
logic by assuming base sorts, function and relation symbols and
axioms, we can define ordered, or categorical theories in \ohol{}
easily by using the two-level type theory structure.
%
We can define an \emph{ordered theory} to be any context $\Gamma$ and
a translation of theories to be a substitution.
%
Then we can define theories for cartesian categories, monoidal
categories, bicartesian categories, strong monads, (strong)
adjunctions.
%
However the utility for applications to programming language semantics
is limited due to the lack of opposite categories.

\subsection{Duality Principle}
One common source of category theoretic arguments is the duality
principle: since the opposite of any category is also a category, any
statement that quantifies over all categories is equally true of their
opposites. Furthermore in a formal language of categories, any
statement can be dualized to a statement where domains and codomains
are swapped and composition is flipped.
%
A similar duality principle holds in SCT.
%
The basic idea is to reverse the orientation of all \emph{set}
expressions by swapping which variable is covariant and which is
contravariant.
%
This has the effect of reversing the order of the context $\Phi$.
%
We can define this as a translation $\cdot^* : SCT \to SCT$ defined on
syntactic forms of SCT in Figure~\ref{fig:duality}.
\begin{figure}
    \begin{align*}
      (\harr P Q)^* &= \harr {P^*} {Q^*}\\
      (\alpha:\cat C)^* &= \alpha: \cat C^*\\
      (\alpha:\cat C; \beta:\cat D)^* &= \beta:\cat D^*; \alpha:\cat C^*\\
      (\Phi,x:R,\Psi)^* &= \Psi^*,x:R^*, \Phi^*\\
      (\homrallXYtoZ \alpha P Q)^* &= \homlallXYtoZ \alpha {P^*} {Q^*}\\
      (\homlallXYtoZ \alpha P Q)^* &= \homrallXYtoZ \alpha {P^*} {Q^*}\\
      (\homrlambdaXatYdotZ x \alpha s)^* &= \homllambdaXatYdotZ x \alpha s^*\\
      (\homrappXtoYatZ s t a)^* &= \homlappXtoYatZ {s^*} {t^*} {a^*}\\
      (\tensorexistsXwithYandZ \beta P Q)^* &=
      \tensorexistsXwithYandZ \beta {Q^*}{P^*}\\
      (\tensorintroatXwithYandZ b s t)^* &= \tensorintroatXwithYandZ {b^*} {t^*} {s^*}\\
      (\punitinXfromYtoZ {\cat C} {a_1}{a_2})^* &= \punitinXfromYtoZ {\cat C^*} {a_1^*}{a_2^*}
    \end{align*}
  \caption{Duality Involution}
  \label{fig:duality}
\end{figure}

\section{Semantics}
Next, we develop the basics of the model theory for \ohol{}. Being a
2-level type theory, \ohol{} has models that are a kind of
\emph{hyperdoctrine}. First, a model must include a model of
intuitionistic type theory to interpret the type-level context
$\Gamma$ and corresponding type/term judgments. Next, there should be
a functor from the category of semantic contexts and substitutions to
a category of models of ordered higher-order logic.

\subsection{Soundness and Completeness}

\subsection{Concrete Models}

Because SCT
includes not just a syntax for category theory, but is also indexed by
an ordinary context of dependent type theory $\Gamma$, a model of SCT
is not given by a \emph{single} virtual double category, but instead
by an \emph{indexed family} of VDCs, parameterized by an object of a
category that interprets dependent type theory. We call such a
structure a \emph{VDC-hyperdoctrine}. We then show how to construct a
from any VDC, a VDC-hyperdoctrine indexed by the category of sets,
which has the intuitive reading that any of the VDC judgments $\Gamma
\vdash J$ is interpreted as a family of $J$-data parameterized
by elements of the set $\Gamma$.

@@

\section{Extended Synthetic Category Theory}
